Plastic in the Urban Freshwater Environment

### 2021-11-11 code/analysis_code/analysis.Rmd
ungroup data for analysis
pt 4 - continue modeling, analysis

# linear regression
```{r}
data <- data %>% select("particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist")

#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit1 <- lm_mod %>% fit(particles_l ~ ., data = data)
#save table
lm_fit1_table <- tidy(lm_fit1)
#view summary of fit
(lm_fit1)
#save file
saveRDS(lm_fit1_table, file = here("results", "lm_fit_table.rds"))
```

@@

#model comparison
comp8_9 <- anova(lm_fit8$fit, lm_fit9$fit, test = "Chisq")
comp8_9
```

@@

Why are there so many rows resulting? When we did the analysis last week, only one row was the result, and it was just rmse and provided a value around 1.

@@

#visualize linear model
t1 <- ggplot(data, aes(x = turbidity.ntu, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. Turbidity", x = "Turbidity (ntu)", y = "Particles/Liter")

#save figure
ggsave(filename = here("results","m_concvturbidity.png"), plot=t1)
```

@@

Here, I am still confused about why so many rmse values are resulting. Earlier, I was getting multiple rmse values but thought it was because I was building a model with all predictors. Here, I am using one predictor but still getting multiple rmse values. If I remember correctly, there was only one rmse value in the exercises we worked on in the previous weeks.

>>>>

# linear regression
```{r}
data <- data %>% select("particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist", "watershed")
#save data subset
saveRDS(data, file = here("data", "processed_data", "data_subset.rds"))
#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit1 <- lm_mod %>% fit(particles_l ~ ., data = data)
#save table
lm_fit1_table <- tidy(lm_fit1)
#view summary of fit
(lm_fit1)
#save file
saveRDS(lm_fit1_table, file = here("results", "lm_fit_table.rds"))
```

@@

#model comparison
comp8_9 <- anova(lm_fit8$fit, lm_fit9$fit, test = "Chisq")
comp8_9

@@

@@

#visualize linear model
t1 <- ggplot(data, aes(x = turbidity.ntu, y = particles_l)) + geom_point() + geom_smooth(method = lm) + labs(title = "Microplastic Concentration vs. Turbidity", x = "Turbidity (ntu)", y = "Particles/Liter")
t1
#save figure
ggsave(filename = here("results","m_concvturbidity.png"), plot=t1)
```

@@

Here, we see that a model with turbidity as the single predictor produces an RMSE of 58. This is significantly better than the full, all-predictor model which had an RMSE of 101.

Going forward, perhaps we could add in additional predictors one at a time to determine what predictors may improve the model performance.

### 2021-11-11 code/analysis_code/modeling.Rmd
pt 4 - continue modeling, analysis

---
title: "land_model"
output: html_document
---

In this portion of the analysis, we will include the land use data for the sample sites to see if the additional predictors can improve our model fit.

@@

#load cleaned land data. 
data <- readRDS(here("data", "processed_data", "athensland.rds"))
data <- data %>% select(-"site", 
                        -"lat",
                        -"long",
                        -"date")
```

@@

RMSE for best LASSO model on the land data is 127.5, compared to 109.4 RMSE for the null model. This tells us our model is not good.

In the perfect outcome/prediction plot, you hope to see data scattered along a 45 degree diagonal line. Instead, data is clustered. However, the scales of the x and y axis are different which may impact how we are viewing the data. Let's visualize with a better scale.

```{r}
#model predictions from tuned model vs actual outcomes SCALE ADJUSTED
lasso_pred_plot_adj <- ggplot(lasso_resid, aes(x = particles_l, y = .pred)) + geom_point() + 
  labs(title = "Predictions vs Actual Outcomes: LASSO", x = "Particles/L Outcome", y = "Particles/L Prediction") + ylim(0, 1000) + xlim(0,1000)
lasso_pred_plot_adj
```

Our outcome vs prediction plot isn't terrible but also isn't great. It looks like predictions are consistently higher than the actual outcomes, since the points fall below the hypothetical 45 degree diagonal line that would represent a perfect fit.

@@

It appears that the addition of land use data does not make our model more informative. In fact, it increases the RMSE to be higher than that of the null model.

>>>>

---
title: "modeling"
output: html_document
---

In this portion of the analysis, we will use the LASSO method for regularization of the model.

@@

#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")

#load cleaned data. 
data <- readRDS(here("data", "processed_data", "mpwrfcombo.rds"))
```

@@

next I want to see what predictors are used in the best lasso model. also need to compare to null model

RMSE for best LASSO model is 100.5, compared to 109.4 RMSE for the null model. Still need to check included predictors for the best LASSO model.

@@

By looking at the predictors used in this fit, we see there are a lot of uninformative predictors included, such as latitude, longitude, site ID, FIPS, etc. Let's look at a subset of predictors to build a model.

Try modeling with data subset, including these variables: "particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist". We'll do this in a separate Rmd document called `subset_modeling.Rmd`

### 2021-11-11 code/analysis_code/subset_modeling.Rmd
pt 4 - continue modeling, analysis

---
title: "land_model"
output: html_document
---

In this portion of the analysis, we will include the land use data for the sample sites to see if the additional predictors can improve our model fit.

@@

#load cleaned land data. 
data <- readRDS(here("data", "processed_data", "athensland.rds"))
data <- data %>% select(-"site", 
                        -"lat",
                        -"long",
                        -"date")
```

@@

RMSE for best LASSO model on the land data is 127.5, compared to 109.4 RMSE for the null model. This tells us our model is not good.

@@

Our outcome vs prediction plot isn't terrible but also isn't great. It looks like predictions are consistently higher than the actual outcomes, since the points fall below the hypothetical 45 degree diagonal line that would represent a perfect fit.

@@

It appears that the addition of land use data does not make our model more informative. In fact, it increases the RMSE to be higher than that of the null model.

>>>>

---
title: "subset_modeling"
output: html_document
---

In this portion of the analysis, we will perform modeling on a smaller subset of data in attempt to improve model performance.

@@

#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")

#load cleaned USBSET data. 
data <- readRDS(here("data", "processed_data", "data_subset.rds"))
```

@@

next I want to see what predictors are used in the best lasso model. also need to compare to null model

RMSE for best LASSO model on the subset data is 113.09, compared to 109.4 RMSE for the null model. This is definitely not a well-performing model, since we want a better RMSE for any model we build than the RMSE for the null model.

@@

Now that we have adjusted the scale, our outcome vs prediction plot looks decent, and there don't appear to be systematic deviations. This indicates that our LASSO model may be flexible enough for our data.

@@

Try modeling with data subset, including these variables: "particles_l", "visual_score", "turbidity.ntu", "temperature.c", "e.coli.cfu", "population", "dist".

The results we get from the initial machine learning modeling and from the modeling of this subset data are not great. Let's think of some ways that we can go about improving our model. 
- An option is variable/predictor removal. We tried that with the subset data. We could attempt further subsets going forward. 
- Variable transformation. Some microplastic studies use log transformation on the concentration values to get a normal distribution. 
- Alternatives: account for different distribution. The distribution of microplastic concentration may be considered more Poisson distributed than normally distributed. First, we can test to see whether it falls in either of these distribution categories, and based on the results we can apply a GLM with Poisson distribution. (Tranformation could make intrepretability more difficult - may want to avoid)
- however, regularization may accomplish our goals more efficiently - we are advised that subset selection is not always a great idea and doesn't neccesarily do the best job of addressing overfitting
- could try ridge regression for alternate form of regularization 
- could try decision tree (though we learned these usually don't perform as well as LASSO)

```{r}
#check for Poisson distribution
#https://stackoverflow.com/questions/59809960/how-do-i-know-if-my-data-fit-a-poisson-distribution-using-r
dispersion_test <- function(x) 
{
  res <- 1-2 * abs((1 - pchisq((sum((x - mean(x))^2)/mean(x)), length(x) - 1))-0.5)

  cat("Dispersion test of count data:\n",
      length(x), " data points.\n",
      "Mean: ",mean(x),"\n",
      "Variance: ",var(x),"\n",
      "Probability of being drawn from Poisson distribution: ", 
      round(res, 3),"\n", sep = "")

  invisible(res)
}

set.seed(1)
x <- data$particles_l

dispersion_test(x)
```

### 2021-11-11 products/manuscript/Manuscript.Rmd
pt 4 - continue modeling, analysis

---
title: "Plastic in the Urban Freshwater Environment"
subtitle: "An Exploratory Study of Microplastics in the Upper Oconee Watershed via Community-Based Efforts"
author: Amelia M. Foley
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
    #reference_docx: style_template.docx #working on creating a template doc with the appropriate style for the conference I am attending next summer (ASTM) so that my manuscript can be automatically formatted
  bookdown::html_document2: 
    toc: false
bibliography: library.bib 
csl: elsevier.csl
---

>>>>

---
title: "Plastic in the Urban Freshwater Environment"
subtitle: "An Exploratory Study of Microplastics in the Upper Oconee Watershed via Community-Based Efforts"
author: Amelia M. Foley
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
    global_numbering: true
    fig_caption: yes
    #reference_docx: style_template.docx #working on creating a template doc with the appropriate style for the conference I am attending next summer (ASTM) so that my manuscript can be automatically formatted
  bookdown::html_document2: 
    toc: false
bibliography: library.bib 
csl: elsevier.csl
---
